---
output: github_document
bibliography: myref_cmp.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# MultRegCMP

MultRegCMP is an R package designed to fit a Bayesian Regression Model for multivariate count data. We assume that the data is distributed according to the Conway-Maxwell-Poisson distribution [@conway1962], and for each response variable, we can associate different covariates. This model allows us to account for correlations between the counts by using latent effects based on the @chib2001 proposal.

You can find more details about this model once our paper is published.

## Installation

You can install the development version of MultRegCMP like so:

``` r
devtools::install_github("mauroflorez/MultRegCMP")
```

### Load package

Here you can see how we can use the package and their functions

```{r loading}
library(MultRegCMP)
## basic example code
```

## Main functions

The function `com_sampler(mu, nu)` allow us to sample data from the CMP with location parameter mu and shape parameter nu. If `ndraws = TRUE` it returns the number of draws needed to sample `n` data. It uses the rejection sampler proposed by @benson2021.

```{r com_sampler}
com_sampler(4, 5)

com_sampler(1, 0.5, 10, ndraws = TRUE)
```

Functions `log_cmp` calculate the log-likelihood of the unnormalized component of the CMP, while `llk_cmp` calculate the approximate log-likelihood of the distribution as suggested by @benson2021. 

```{r llk}
llk_cmp(y = 2, mu = 2, nu = 1.5)
```

Finally, to implement our model we use the main function `mcmc_cmp`. This function requires two important componentes; `X` is the covariates information, this object is a list, where every element of the list is a covariance matrix for every response. On the other hand, `y` is a matrix (nxJ) containing all the J response variables and the n observations. We can see an example below

```{r data}
n = 50; J = 2
X = list(matrix(rnorm(3*n), ncol = 3), matrix(rnorm(3*n), ncol = 3))
beta <- list(c(1,0.1, 1), c(0, 0.5, -0.5))
mu <- exp(prod_list(X, beta))
y = matrix(rpois(n = length(mu), lambda = mu), nrow = n)
```

To fit our model, we require the number of draws `S`, the number of samples burn `nburn` and some parameters for the MCMC - Exchange Algorithm steps.

```{r model}
fit <- mcmc_cmp(y, X, S = 40000, nburn = 1000, v0 = 40, scale_cov_b = 0.8, scale_cov_beta = 0.04, scale_cov_gamma = 0.06, progress = "bar")
```
### Parameter Estimation

We can see the estimations of the model as

```{r estimation}
fit$estimation_beta
fit$estimation_gamma
```
Also, we plot the trace plots of the parameters associated to the location and shape parameters to check convergence visually. Here we can see the trace plots associated to the first response variable:

```{r tp, echo=FALSE, fig.height=6, fig.width=6, out.width='50%'}
post_beta <- fit$posterior_beta[[1]]
post_gamma <- fit$posterior_gamma[[1]]

plot(NULL, ylim = c(min(post_beta, post_gamma) - 0.2, max(post_beta, post_gamma) + 0.2),
     xlim = c(0, nrow(post_beta)),
     main = "Trace plots - Exchange Algorithm", xlab = "Iteration", ylab = "Posterior")
     
for(j in 1:ncol(post_beta)){
  lines(post_beta[,j], lty = j)
  lines(post_gamma[,j], col = "darkgoldenrod", lty = j)
}

legend("top", legend = c(expression(beta[11]), expression(beta[12]), expression(beta[13])), 
       lty = c(1,2,3), horiz = TRUE)
legend("bottom", legend = c(expression(gamma[11]), expression(gamma[12]), expression(gamma[13])), 
       col = rep("darkgoldenrod", 3), lty = c(1,2,3), horiz = TRUE)
```

and similarly for the parameters of the second response.

```{r tp2, echo=FALSE, fig.height=6, fig.width=6, out.width='50%'}
post_beta <- fit$posterior_beta[[2]]
post_gamma <- fit$posterior_gamma[[2]]

plot(NULL, ylim = c(min(post_beta, post_gamma) - 0.2, max(post_beta, post_gamma) + 0.2),
     xlim = c(0, nrow(post_beta)),
     main = "Trace plots - Exchange Algorithm", xlab = "Iteration", ylab = "Posterior")
     
for(j in 1:ncol(post_beta)){
  lines(post_beta[,j], lty = j)
  lines(post_gamma[,j], col = "darkgoldenrod", lty = j)
}

legend("top", legend = c(expression(beta[21]), expression(beta[22]), expression(beta[23])), 
       lty = c(1,2,3), horiz = TRUE)
legend("bottom", legend = c(expression(gamma[21]), expression(gamma[22]), expression(gamma[23])), 
       col = rep("darkgoldenrod", 3), lty = c(1,2,3), horiz = TRUE)
```

Additionally `DIC_cmp` allows to calculate an approximation of the Deviance Information Criterion [@spiegelhalterbayesian] of the fitted model for each response. 

```{r DIC}
DIC_cmp(fit)
```
## References
