---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# MultRegCMP

MultRegCMP is an R package designed to fit a Bayesian Regression Model for multivariate count data. We assume that the data is distributed according to the Conway-Maxwell-Poisson distribution, and for each response variable, we can associate different covariates. This model allows us to account for correlations between the counts.

You can find more details about this model once our paper is published.

## Installation

You can install the development version of MultRegCMP like so:

``` r
devtools::install_github("mauroflorez/MultRegCMP")
```

## Example

Here you can see how we can use the package and their functions

```{r loading}
library(MultRegCMP)
## basic example code
```

The function `com_sampler(mu, nu)` allow us to sample data from the CMP with location parameter mu and shape parameter nu. If `ndraws = TRUE` it returns the number of draws needed to sample `n` data.

```{r com_sampler}
com_sampler(4, 5)

com_sampler(1, 0.5, 10, ndraws = TRUE)
```

Functions `log_cmp` calculate the log-likelihood of the unnormalized component of the CMP, whilea `llk_cmp` calculate the approximate log-likelihood of the distribution as suggested by Benson and Friel (2021). 

```{r llk}
llk_cmp(y = 2, mu = 2, nu = 1.5)
```

Finally, to implement our model we use the main function `mcmc_cmp`. This function requires two important componentes; `X` is the covariates information, this object is a list, where every element of the list is a covariance matrix for every response. On the other hand, `y` is a matrix (nxJ) containing all the J response variables and the n observations. We can see an example below

```{r data}
n = 50; J = 2
X = list(matrix(rnorm(3*n), ncol = 3), matrix(rnorm(3*n), ncol = 3))
beta <- list(c(1,0.1, 1), c(0, 0.5, -0.5))
mu <- exp(prod_list(X, beta))
y = matrix(rpois(n = length(mu), lambda = mu), nrow = n)
```

To fit our model, we require the number of draws `S`, the number of samples burn `nburn` and some parameters for the MCMC - Exchange Algorithm steps.

```{r model}
fit <- mcmc_cmp(y, X, S = 50000, nburn = 1000, scale_cov_b = 0.8, scale_cov_beta = 0.04, scale_cov_gamma = 0.06, progress = "bar")
```

We can see the estimations of the model as

```{r estimation}
fit$estimation_beta
fit$estimation_gamma
```

or see the trace plots

```{r tp, fig.height=4, fig.width=6}
plot.ts(fit$posterior_beta[[1]], main = "Traceplot - Exchange Algorithm", ylab = "Beta_1")
plot.ts(fit$posterior_beta[[2]], main = "Traceplot - Exchange Algorithm", ylab = "Beta_2")
```

and for the shape parameters

```{r tps, fig.height=4, fig.width=6}
plot.ts(fit$posterior_gamma[[1]], main = "Traceplot - Exchange Algorithm", ylab = "Gamma_1")
plot.ts(fit$posterior_gamma[[2]], main = "Traceplot - Exchange Algorithm", ylab = "Gamma_2")
```

Additionally `DIC_cmp` allows to calculate an approximation of the Deviance Information Criterion of the fitted model for each response.

```{r DIC}
DIC_cmp(fit, X, y)
```

We can check the fitting of the model to the real data using the function `fitting_plot`

```{r plot}
fitting_plots(fit, type = "bar", S = 1000)
```
