---
title: "Example Real Data: English Premier League - Season 20-21"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Setup

```{r setup}
library(MultRegCMP)
library(ggplot2)
library(ggrepel)
library(latex2exp)
```

The data contained in the package `epl_20_21` corresponds to the scores of the games of the English Premier League (EPL) during the season 2020-2021. The data can be accessed as:

```{r }
data("epl_20_21")
```

```{r }
head(epl_20_21)
```

## Fitting

We can fit a bivariate regression model for the goals scored by the home and away team, as stated in the paper (in revision).

We first adjust the data to use the `mcmc_cmp` function

```{r }
y <- cbind(epl_20_21$HG, epl_20_21$AG)

X1 <- model.matrix(HG ~ 1 + HomeTeam + AwayTeam, contrasts = list(HomeTeam = "contr.sum", AwayTeam = "contr.sum"), data = epl_20_21)

X2 <- model.matrix(AG ~ 1 + AwayTeam + HomeTeam, contrasts = list(AwayTeam = "contr.sum", HomeTeam = "contr.sum"), data = epl_20_21)

X <- list(X1, X2)
```

Then, we fit the model as

```{r }
fit <- mcmc_cmp(y = y, X = X, S = 100000, nburn = 10000, scale_cov_b = 0.6, 
                scale_cov_beta = 0.01, scale_cov_gamma = 0.02, progress = "bar",
                v0 = 50)
```
## Convergence Diagnostic

After fitting the model, we can check the convergence visually of the parameters by plotting their trace plots.

```{r fig.height=4, fig.width=6}
list_colors <- c("#D1D1D1", "#BDBDBD", "#333333", "#424242", "#575757",
                 "#6B6B6B", "#000000", "#A8A8A8", "#949494", "#808080")
list_colors <- rep(list_colors, 10)[1:ncol(fit$posterior_beta[[1]])]

plot(NULL, ylim = c(min(fit$posterior_beta[[1]]), max(fit$posterior_beta[[1]])), xlim = c(0, nrow(fit$posterior_beta[[1]])),
     ylab = expression(beta[ij]), xlab = "Iteration",
     main = "Traceplot of the parameters associated to the Home Goals")

for(beta in 1:ncol(fit$posterior_beta[[1]])){
  lines(fit$posterior_beta[[1]][,beta], col = list_colors[beta])
}

```

We can check the fitting of the model to the real data using the function `fitting_plot` to plot the Rootograms for each response.

```{r plot, echo=FALSE, fig.height=3, fig.width=5}
fitting_plots(fit, type = "bar", S = 1000)
```

## Parameters Estimation

We can also plot the estimated parameters to facilitate their interpretation.

```{r }
beta1_est <- fit$estimation_beta[[1]]
beta2_est <- fit$estimation_beta[[2]]

off_home <- beta1_est[2:20]
off_home <- c(off_home, 0 - sum(off_home)) #Sum-to-zero constraint
def_away <- beta1_est[21:39]
def_away <- c(def_away, 0 - sum(def_away)) #Sum-to-zero constraint

def_home <- beta2_est[21:39]
def_home <- c(def_home, 0 - sum(def_home)) #Sum-to-zero constraint
off_away <- beta2_est[2:20]
off_away <- c(off_away, 0 - sum(off_away)) #Sum-to-zero constraint
```

```{r echo=FALSE}
personal_theme =   theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

table_teams = data.frame("Home" = levels(factor(epl_20_21$HomeTeam)), 
                         "Attack Home" = exp(off_home), "Defense Home" = exp(-def_home), 
                         "Attack Away" = exp(off_away), "Defense Away" = exp(-def_away))
```

Below we can find the plots for the strengths playing at Home and Away. With these plots we can identify the teams that had a good season and those who had the worst performance.

```{r parameters, echo=TRUE, fig.height=4, fig.width=6, message=FALSE, warning=FALSE, out.width = '50%'}
ggplot(table_teams, aes(x = Attack.Home, y = Defense.Home)) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "grey50") +
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50") +
  geom_point() +
  geom_label_repel(aes(label = Home), box.padding = 0.25, point.padding = 0.5, 
                   segment.color = "grey50") +
  labs(title = "Strengths playing Home",
       x = TeX("Attack $(\\exp(\\beta^{\\omega_{H_i}}))$"), 
       y = TeX("Defense $(\\exp(-\\beta^{\\delta_{H_i}}))$")) + 
  personal_theme

ggplot(table_teams, aes(x = Attack.Away, y = Defense.Away)) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "grey50") +
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50") +
  geom_point() +
  geom_label_repel(aes(label = Home), box.padding = 0.25, point.padding = 0.5,
                   segment.color = "grey50") +
  labs(title = "Strengths playing Away",
       x = TeX("Attack $(\\exp(\\beta^{\\omega_{A_i}}))$"), 
       y = TeX("Defense $(\\exp(-\\beta^{\\delta_{A_i}}))$")) + 
  personal_theme
```
